---
title: "Time Series Analysis with Window Functions"
author: "Guibril Ramde"
date: "2026-02-12"
format: html
execute:
  echo: true
  warning: false
  message: false
---

## ##Overview

This analysis uses a synthetic datasets containing daily cryptocurrency prices for Bitcoin and Ethereum beginning January 1, 2022.

The objectives are:

-   Compute the Year-to-Date (YTD) average price
-   Compute the six-day moving average
-   Demonstrate the use of window functions in `dplyr`

## ##Running Code

Generating the time series dataset

```{r}
library(dplyr)
library(lubridate)
library(tidyr)
library(zoo)

set.seed(123)
dates <- seq.Date(from = as.Date("2022-01-01"), 
                  to = as.Date("2022-12-31"), by = "day")
crypto <- c("Bitcoin", "Ethereum")
data <- expand.grid(date = dates, crypto = crypto)

# Simulate daily closing prices
data$price <- round(runif(nrow(data), 1000, 50000), 2)

# Preview dataset
head(data)

```

Applying window Functions to analyze the data

```{r}
data_analysis <- data %>%
  arrange(crypto, date) %>%
  group_by(crypto) %>%
  
  mutate(ytd_avg = cummean(price)) %>%
  
  mutate( ma_6day = zoo::rollmean(price, k = 6, fill = NA, align = "right")) %>%
  ungroup()
  

head(data_analysis, 10)  
```

**##Conclusion**

This analysis demonstrates how window functions allow cumulative and rolling\

calculations without collapsing the datasets. These techniques are widely used\

in financial modeling, time series analysis, and predictive analytics.

```         
```
